---
title: "Exploratory Analysis 1"
author: "Sarah Hensley"
date: "2024-02-01"
output: html_document
---

```{r, echo=FALSE, message=FALSE}
# loading libraries
library(mgcv)
library(here)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(readr)

# loading in the data
here()
setwd("C:/Users/hensl/Downloads/NOAA IN FISH! Internship/IN FISH Woods Hole Project")
data_comb <- read.csv("GAM Data/updated cb ratios.csv")

# renaming species codes
data_comb$svspp[data_comb$svspp == 15] <- 'Spiny.Dogfish'
data_comb$svspp[data_comb$svspp == 23] <- 'Winter.Skate'
data_comb$svspp[data_comb$svspp == 75] <- 'Pollock'
data_comb$svspp[data_comb$svspp == 76] <- 'White.Hake'
data_comb$svspp[data_comb$svspp == 77] <- 'Red.Hake'
data_comb$svspp[data_comb$svspp == 197] <- 'Goosefish'
data_comb$svspp[data_comb$svspp == 28] <- 'Thorny.Skate'
data_comb$svspp[data_comb$svspp == 73] <- 'Atlantic.Cod'
data_comb$svspp[data_comb$svspp == 164] <- 'Sea.Raven'
data_comb$svspp[data_comb$svspp == 72] <- 'Silver.Hake'
data_comb$svspp[data_comb$svspp == 103] <- 'Summer.Flounder'
data_comb$svspp[data_comb$svspp == 104] <- 'Fourspot.Flounder'
data_comb$svspp[data_comb$svspp == 108] <- 'Windowpane'
data_comb$svspp[data_comb$svspp == 112] <- 'Buckler.Dory'
data_comb$svspp[data_comb$svspp == 135] <- 'Blue.Fish'
data_comb$svspp[data_comb$svspp == 163] <- 'Longhorn.Sculpin'
data_comb$svspp[data_comb$svspp == 172] <- 'Striped.Searobin'

# omitting all NA's
data_comb_f <- data_comb %>% na.omit(cb)
```

Filtering through the data as requested: transforming c:b ratios on a log scale, starting time series in 1981, and removing 0 c:b values

```{r}
# omitting all 0's from the dataset
rows_to_keep <- data_comb_f$cb > 0.000000000
data_comb_f2 <- data_comb_f[rows_to_keep, ]

# starting time series in 1981 for sensitivity analysis
data_comb_f2 <- data_comb_f %>% 
  filter(year > 1980)

# log transforming c:b ratios 
data_comb_f2$cb <- log(data_comb_f2$cb)
```

FILTERED VERSION: Breaking the above plots into a legible format by species, to assess species specific prey consumption over time

```{r, out.width='100%', fig.height=6, echo=FALSE}
# Get unique levels of svspp
svspp_levels <- unique(data_comb_f2$svspp)

# Create a list to store plots
dist_plots_list <- list()

# Loop through svspp levels
for (svspp_level in svspp_levels) {
  # Filter data for the current svspp level
  filtered_data <- data_comb_f2 %>%
    filter(svspp == svspp_level)

  # Create the plot
  dist_plot <- ggplot(filtered_data, aes(x = year, y = cb, fill = prey)) +
    geom_point(stat = 'identity') +
    geom_smooth(method = lm, aes(color = prey), linewidth = 1, se = FALSE) +
    xlab("Time (Years)") +
    ylab("log(Consumption:Biomass Ratios)") + 
    facet_wrap(~prey + seacat, scales = "free") +
    scale_x_continuous(breaks = c(1980, 2000, 2020)) +
    theme_bw() +
    theme(plot.title = element_text(size = 15)) +
    theme(axis.text.x = element_text(size = 10), 
          axis.text.y = element_text(size = 10),
          axis.title.y = element_text(size = 13),
          axis.title.x = element_text(size = 13)) +
    ggtitle(paste("Consumption:Biomass Over Time for", svspp_level))

  # Add the plot to the list
  dist_plots_list[[svspp_level]] <- dist_plot
}

# Print the plots
for (dist_plot in dist_plots_list) {
  print(dist_plot)
}

```

FILTERED VERSION: Doing this same kind of analysis but looking at how the values compare with bottom temperature

```{r, out.width='100%', fig.height=7, echo=FALSE}
# Get unique levels of svspp
svspp_levels <- unique(data_comb_f2$svspp)

# Create a list to store plots
dist_plots_list <- list()

# Loop through svspp levels
for (svspp_level in svspp_levels) {
  # Filter data for the current svspp level
  filtered_data <- data_comb_f2 %>%
    filter(svspp == svspp_level)

  # Create the plot
  dist_plot <- ggplot(filtered_data, aes(x = bt, y = cb, fill = prey)) +
    geom_point(stat = 'identity') +
    geom_smooth(method = lm, aes(color = prey), linewidth = 1, se = FALSE) +
    xlab("Bottom Temperature (C)") +
    ylab("log(Consumption:Biomass Ratios)") + 
    facet_wrap(~prey + seacat, scales = "free") +
    theme_bw() +
    theme(plot.title = element_text(size = 15)) +
    theme(axis.text.x = element_text(size = 10), 
          axis.text.y = element_text(size = 10),
          axis.title.y = element_text(size = 13),
          axis.title.x = element_text(size = 13)) +
    ggtitle(paste("Consumption:Biomass v. Bottom Temperature for", svspp_level))

  # Add the plot to the list
  dist_plots_list[[svspp_level]] <- dist_plot
}

# Print the plots
for (dist_plot in dist_plots_list) {
  print(dist_plot)
}

```

## Bottom Temperature Overview
```{r, out.width='100%', fig.height=6, echo=FALSE}
dist_plot <- ggplot(data_comb_f2, aes(x = year, y = bt, fill = prey)) +
    geom_point(stat = 'identity') +
    geom_smooth(method = lm, aes(color = prey), linewidth = 1, se = FALSE) +
    xlab("Time (Years)") +
    ylab("Bottom Temperature (C)") + 
    facet_wrap(~prey + seacat, scales = "free") +
    theme_bw() +
    theme(plot.title = element_text(size = 15)) +
    theme(axis.text.x = element_text(size = 10), 
          axis.text.y = element_text(size = 10),
          axis.title.y = element_text(size = 13),
          axis.title.x = element_text(size = 13)) +
    ggtitle(paste("Bottom Temperature Through Time"))
dist_plot
```



