---
title: "Exploratory Analysis for C:B Data"
author: "Sarah Hensley"
date: "2024-01-23"
output:
  html_document: default
---

# Consumption to Biomass Data - Exploratory Analysis

This dataset covers 17 different fish species that inhabit the Northwest Atlantic. Prey consumption is displayed as a ratio to biomass for the respective species, and is broken down by prey type. Data covers a rough time frame from 1971-2021.


```{r, echo=FALSE, message=FALSE}
# loading libraries
library(mgcv)
library(here)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(readr)
library(patchwork)

# loading in the data
here()
data_comb <- read.csv("data/updated cb ratios.csv")

# renaming species codes
data_comb$svspp[data_comb$svspp == 15] <- 'Spiny.Dogfish'
data_comb$svspp[data_comb$svspp == 23] <- 'Winter.Skate'
data_comb$svspp[data_comb$svspp == 75] <- 'Pollock'
data_comb$svspp[data_comb$svspp == 76] <- 'White.Hake'
data_comb$svspp[data_comb$svspp == 77] <- 'Red.Hake'
data_comb$svspp[data_comb$svspp == 197] <- 'Goosefish'
data_comb$svspp[data_comb$svspp == 28] <- 'Thorny.Skate'
data_comb$svspp[data_comb$svspp == 73] <- 'Atlantic.Cod'
data_comb$svspp[data_comb$svspp == 164] <- 'Sea.Raven'
data_comb$svspp[data_comb$svspp == 72] <- 'Silver.Hake'
data_comb$svspp[data_comb$svspp == 103] <- 'Summer.Flounder'
data_comb$svspp[data_comb$svspp == 104] <- 'Fourspot.Flounder'
data_comb$svspp[data_comb$svspp == 108] <- 'Windowpane'
data_comb$svspp[data_comb$svspp == 112] <- 'Buckler.Dory'
data_comb$svspp[data_comb$svspp == 135] <- 'Blue.Fish'
data_comb$svspp[data_comb$svspp == 163] <- 'Longhorn.Sculpin'
data_comb$svspp[data_comb$svspp == 172] <- 'Striped.Searobin'

# omitting all NA's
data_comb_f <- data_comb %>% na.omit(cb)

# omitting true zeros
data_comb_f <- data_comb_f %>% 
  filter(CMT > 0.000000e+00)
```

### Looking at Consumption vs Biomass Over Time Across Predator, Prey and Seasons
```{r, out.width='100%', fig.height=6, echo=FALSE}
# Create a list to store plots
dist_plots_list <- list()

# Get unique levels of svspp
svspp_levels <- unique(data_comb_f$svspp)

# Loop through svspp levels
for (svspp_level in svspp_levels) {
  # Filter data for the current svspp level
  filtered_data <- data_comb_f %>%
    filter(svspp == svspp_level)

  # Create the consumption plot
  dist_plot_consumption <- ggplot(filtered_data, aes(x = year, y = CMT)) +
    geom_point(stat = 'identity') +
    geom_smooth(method = lm, aes(color = prey), linewidth = 1, se = FALSE) +
    xlab("Time (year)") +
    ylab("Consumption (metric tons)") + 
    facet_wrap(~prey + seacat, scales = "free") +
    scale_x_continuous(breaks = c(1980, 2000, 2020)) +
    theme(legend.position = "none") +
    guides(color = "none") +
    theme_bw() +
    theme(plot.title = element_text(size = 15)) +
    theme(axis.text.x = element_text(size = 10), 
          axis.text.y = element_text(size = 10),
          axis.title.y = element_text(size = 13),
          axis.title.x = element_text(size = 13)) +
    ggtitle(paste("Consumption:", svspp_level))
  
  # Create the biomass plot
  dist_plot_biomass <- ggplot(filtered_data, aes(x = year, y = biompop, fill = prey)) +
    geom_point(stat = 'identity') +
    geom_smooth(method = lm, aes(color = prey), linewidth = 1, se = FALSE) +
    xlab("Time (year)") +
    ylab("Biomass (kg)") + 
    facet_wrap(~prey + seacat, scales = "free") +
    theme_bw() +
    theme(plot.title = element_text(size = 15)) +
    theme(axis.text.x = element_text(size = 10), 
          axis.text.y = element_text(size = 10),
          axis.title.y = element_text(size = 13),
          axis.title.x = element_text(size = 13)) +
    ggtitle(paste("Biomass:", svspp_level))

  # Add the plot to the list
  dist_plots_list[[svspp_level]] <- dist_plot_consumption + dist_plot_biomass
}

# Print the plots
for (dist_plot in dist_plots_list) {
  print(dist_plot)
}

```

### Doing the same thing as above but just grouping it by all prey to look at total consumption and biomass over time side by side

```{r, out.width='100%', fig.height=6, echo=FALSE}
library(patchwork)

# creating list to save plots
dist_plots_list <- list()

# Get unique levels of svspp
svspp_levels <- unique(data_comb_f$svspp)

# Loop through svspp levels
for (svspp_level in svspp_levels) {
  # Filter data for the current svspp level
  filtered_data <- data_comb_f %>%
    filter(svspp == svspp_level & prey == "all")

  # Create the consumption plot
  dist_plot_consumption <- ggplot(filtered_data, aes(x = year, y = CMT)) +
    geom_point(stat = 'identity') +
    geom_smooth(method = lm, linewidth = 1, se = FALSE) +
    xlab("Time (year)") +
    ylab("Consumption (metric tons)") + 
    facet_wrap(~seacat, scales = "free") +
    scale_x_continuous(breaks = c(1980, 2000, 2020)) +
    theme(legend.text = element_text(size = 5)) +
    theme_bw() +
    theme(plot.title = element_text(size = 15)) +
    theme(axis.text.x = element_text(size = 10), 
          axis.text.y = element_text(size = 10),
          axis.title.y = element_text(size = 13),
          axis.title.x = element_text(size = 13)) +
    ggtitle(paste("Consumption:", svspp_level))
  
   # Create the biomass plot
  dist_plot_biomass <- ggplot(filtered_data, aes(x = year, y = biompop)) +
    geom_point(stat = 'identity') +
    geom_smooth(method = lm, linewidth = 1, se = FALSE) +
    xlab("Time (year)") +
    ylab("Biomass (kg)") + 
    facet_wrap(~seacat, scales = "free") +
    theme_bw() +
    theme(plot.title = element_text(size = 15)) +
    theme(axis.text.x = element_text(size = 10), 
          axis.text.y = element_text(size = 10),
          axis.title.y = element_text(size = 13),
          axis.title.x = element_text(size = 13)) +
    ggtitle(paste("Biomass:", svspp_level))

  # Add the plot to the list
  dist_plots_list[[svspp_level]] <- dist_plot_consumption + dist_plot_biomass
}

# Print the plots
for (dist_plot in dist_plots_list) {
  print(dist_plot)
}

```

#### Questions Regarding these Outputs
It appears as if the results prior to 1981 are showing both significantly more biomass and consumption; do we need to account for this or because they are both changing in the same direction we're fine?
